---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.path = "man/figures/README-",
  fig.alt = "Depth-velocity diagram of the eight default mesohabitat classes"
)
library(hydromeso)
```

# hydromeso

<!-- badges: start -->
Development status: version 0.1.0; not yet on CRAN.
<!-- badges: end -->

`hydromeso` classifies water depth and velocity into nominal fluvial hydraulic
mesohabitat categories. It supports numeric values, ordinary tables,
`terra::SpatVector` features, rasters, multiple hydraulic scenarios, and
validated custom rectangular schemes. Its only non-base runtime dependency is
`terra`.

## Installation

```r
# From a local source checkout:
install.packages("hydromeso", repos = NULL, type = "source")
```

## Default classification

The authoritative default uses metres and metres per second. Lower bounds are
inclusive and upper bounds are exclusive.

|Class|Mesohabitat|Depth (m)|Velocity (m/s)|
|---:|:---|:---|:---|
|1|Shallow Pool|< 0.61|< 0.30|
|2|Medium Pool|>= 0.61 and < 1.37|< 0.30|
|3|Deep Pool|>= 1.37|< 0.30|
|4|Slow Riffle|< 0.61|>= 0.30 and < 0.61|
|5|Fast Riffle|< 0.61|>= 0.61|
|6|Raceway|>= 0.61 and < 1.37|>= 0.30 and < 0.61|
|7|Faster than Raceway|>= 0.61 and < 1.37|>= 0.61|
|8|Faster than Deep Pool|>= 1.37|>= 0.30|

```{r scheme-plot, fig.width=7, fig.height=5}
plot_meso_scheme()
```

## Quick examples

```{r values}
classify_mesohabitat_values(
  depth = c(0, 0.61, 1.37, NA),
  velocity = c(0, 0.30, 0.61, 0.2)
)
```

```{r table}
classified <- classify_mesohabitat_table(
  hydromeso_example, "depth", "velocity"
)
head(classified)
summarize_mesohabitat(classified)
```

```{r vector}
points <- mesohabitat_example_vector()
classified_points <- classify_mesohabitat_vector(points, "depth", "velocity")
classified_points
```

```{r raster}
hydraulics <- mesohabitat_example_rasters()
classes <- classify_mesohabitat_raster(hydraulics$depth, hydraulics$velocity)
classes
summarize_mesohabitat(classes)
```

Custom schemes use the same engine:

```{r custom}
rules <- data.frame(
  class_id = c(1L, 2L), label = c("Slow", "Fast"),
  depth_min = c(0, 0), depth_max = c(Inf, Inf),
  velocity_min = c(0, 0.5), velocity_max = c(0.5, Inf)
)
custom <- meso_scheme(rules, name = "Two velocity classes")
classify_mesohabitat_values(c(0.2, 1), c(0.2, 0.8), custom)
```

## Interpretation and citation

The output is a depth-velocity hydraulic classification, not proof of habitat
quality or fish occupancy. Interpretation depends on species and life stage,
stream type, substrate, cover, connectivity, water quality, temperature, flow
regime, hydraulic-model resolution, and local validation. Class IDs are labels,
not ordinal scores. Users must decide how dry cells are represented in their
own HEC-RAS or other model outputs; `dry_threshold = NULL` preserves the exact
scheme and classifies zero depth and zero velocity as Class 1.

The default eight-class table is from Cordero and Harris (Preprint),
*Semi-Supervised and Supervised Machine Learning Approaches to Predicting
Fluvial Mesohabitats from Satellite Data*, [doi:10.2139/ssrn.7100727](https://doi.org/10.2139/ssrn.7100727).
The broader framework is informed by Aadland (1993), *Stream Habitat Types:
Their Fish Assemblages and Relationship to Flow*,
[doi:10.1577/1548-8675(1993)013<0790:SHTTFA>2.3.CO;2](https://doi.org/10.1577/1548-8675%281993%29013%3C0790%3ASHTTFA%3E2.3.CO%3B2).

Source code and issue tracking are available at
[github.com/el-cordero/hydromeso](https://github.com/el-cordero/hydromeso).
