---
title: "HEC-RAS depth and velocity GeoTIFF workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HEC-RAS depth and velocity GeoTIFF workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library(hydromeso)
```

HEC-RAS workflows commonly begin with paired depth and velocity GeoTIFFs.
The package includes a compact pair derived from the 31 May 2022 HEC-RAS
model output used by Cordero and Harris (Preprint). These are real hydraulic
surfaces from the confluence of the Big Blue and Kansas Rivers near Manhattan,
Kansas, rather than synthetic cells.

The source rasters were cropped to the detailed manuscript-map extent,
aggregated from 3-foot to 18-foot cells, and converted from feet and feet per
second to metres and metres per second. The raster coordinates remain in the
source project's documented NAD 1983 (CORS96) StatePlane Kansas North system.

## Inspect the hydraulic inputs

The helper can return either installed GeoTIFF paths or loaded rasters:

```{r hecras-inputs, fig.width=10, fig.height=4.8, fig.bg="white", fig.cap="Depth and velocity surfaces derived from the 31 May 2022 HEC-RAS output at the Big Blue-Kansas Rivers confluence."}
paths <- mesohabitat_example_rasters(paths = TRUE)
basename(unlist(paths))

h <- mesohabitat_example_rasters()
terra::compareGeom(h$depth, h$velocity)

old_par <- graphics::par(mfrow = c(1, 2), mar = c(1, 1, 3, 4))
terra::plot(h$depth, col = grDevices::hcl.colors(40, "Blues 3"),
            main = "Depth (m)", axes = FALSE)
terra::plot(h$velocity, col = grDevices::hcl.colors(40, "YlOrRd"),
            main = "Velocity (m/s)", axes = FALSE)
graphics::par(old_par)
```

Inspect units, dry-cell encoding, CRS, extent, resolution, origin, and
dimensions before classifying any HEC-RAS export. The example is already in the
metric hydraulic units required by the default scheme.

## Classify the river

```{r hecras-classification, fig.width=10, fig.height=6.2, fig.bg="white", fig.cap="Eight-class hydraulic mesohabitats derived from the real HEC-RAS sample."}
classes <- classify_mesohabitat_raster(h$depth, h$velocity)
class_colors <- mesohabitat_palette()
plot_mesohabitat(
  classes,
  main = "Big Blue-Kansas Rivers mesohabitats — 31 May 2022",
  axes = FALSE,
  palette = class_colors,
  plg = list(x = "bottom", ncol = 4, cex = 0.75)
)
```

The white background is outside the HEC-RAS wetted surface. Class IDs describe
depth-velocity combinations; they are not habitat-quality scores.

## Work with an area of interest

An AOI can be a `SpatVector` or a supported vector path. By default both crop
and polygon mask are applied. This reproducible rectangle selects the central
part of the included river sample:

```{r hecras-aoi, fig.width=10, fig.height=5.2, fig.bg="white", fig.cap="Mesohabitat output cropped to a central area of interest."}
e <- terra::ext(h$depth)
dx <- terra::xmax(e) - terra::xmin(e)
dy <- terra::ymax(e) - terra::ymin(e)
aoi <- terra::as.polygons(
  terra::ext(
    terra::xmin(e) + 0.18 * dx, terra::xmax(e) - 0.18 * dx,
    terra::ymin(e) + 0.12 * dy, terra::ymax(e) - 0.12 * dy
  ),
  crs = terra::crs(h$depth)
)
cropped <- classify_mesohabitat_raster(h$depth, h$velocity, aoi = aoi)
plot_mesohabitat(
  cropped, axes = FALSE, palette = class_colors,
  plg = list(x = "bottom", ncol = 4, cex = 0.75)
)
```

When input geometry differs, `align = "velocity_to_depth"` uses bilinear
interpolation to project or resample continuous velocity to the depth template.
Assigning a known missing CRS with `assume_crs` describes existing coordinates;
it does not move them.

## Process and export scenarios

For multiple dates, use unique matching names rather than relying on file
order. Output can be written to a GeoTIFF with an accompanying class-table CSV:

```{r hecras-export}
series <- classify_mesohabitat_series(
  list(`2022-05-31` = h$depth),
  list(`2022-05-31` = h$velocity)
)
out <- file.path(tempdir(), "hecras_mesohabitat.tif")
write_mesohabitat(series[[1]], out, sidecar = TRUE, overwrite = TRUE)
```

Determine how dry HEC-RAS cells are represented before choosing
`dry_threshold`; the package does not silently redefine zero depth.
