---
title: "HEC-RAS depth and velocity GeoTIFF workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HEC-RAS depth and velocity GeoTIFF workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library(hydromeso)
```

HEC-RAS workflows commonly begin with paired depth and velocity GeoTIFFs.
Inspect their units, dry-cell encoding, CRS, extent, resolution, origin, and
dimensions before classification. The following self-contained example stands
in for file-backed inputs:

```{r}
h <- mesohabitat_example_rasters()
terra::compareGeom(h$depth, h$velocity)
classes <- classify_mesohabitat_raster(h$depth, h$velocity)
```

Assigning a known missing CRS describes existing coordinates; it does not move
them. `assume_crs` makes that assignment explicit and reports it. Projecting
coordinates is different. When geometry differs,
`align = "velocity_to_depth"` uses bilinear interpolation to project or
resample continuous velocity directly to the depth template.

An AOI can be a `SpatVector` or supported path. By default both crop and polygon
mask are applied:

```{r}
aoi <- terra::as.polygons(terra::ext(500000, 500020, 4400000, 4400020),
                          crs = terra::crs(h$depth))
cropped <- classify_mesohabitat_raster(h$depth, h$velocity, aoi = aoi)
```

For multiple dates, use unique matching names rather than relying on file
order. Output can be written to a GeoTIFF and accompanied by a class-table CSV:

```{r}
series <- classify_mesohabitat_series(list(example = h$depth),
                                      list(example = h$velocity))
out <- file.path(tempdir(), "hecras_mesohabitat.tif")
write_mesohabitat(series[[1]], out, overwrite = TRUE)
```

Determine how dry HEC-RAS cells are represented before choosing
`dry_threshold`; the package does not silently redefine zero depth.
